---
title: "Explained: Principal Component Analysis"
output: html_notebook
---

# {.tabset}

```{r setup, include=FALSE}
library(caret)
library(dplyr)
library(ggplot2)
library(gridExtra)  # Side by side plots
library(hurwitzLab) # Color palettes
library(MethComp)   # Deming regression
library(parsnip)
library(purrr)      #map
library(stringr)
library(tibble)
library(tidyr)

source("helpers.R")

theme_set(theme_light() +
          theme(plot.title = element_text(hjust=0.5),
                plot.subtitle = element_text(hjust=0.5)))
```

## Intoduction

### The Data

This document is intended to give an intuitive sense of how PCA works.

I will be using the mtcars dataset built-in to R. For our purposes, we will only
be looking at 3 variables: weight (`wt`), horsepower (`hp`), and mpg. Prior to
looking at the data, I will split into training and testing sets. The training set
has 80% of the data. Here is what they look like.

```{r data_select, echo=FALSE}
df <- mtcars %>% select(mpg, wt, hp)

set.seed(3456)
trainIndex <- createDataPartition(df$mpg, p = .8,
                                  list = FALSE,
                                  times = 1)
train <- df[trainIndex,]
test <- df[-trainIndex,]

train
```

### Sample Plot

```{r raw_plot, echo=FALSE}
train %>% ggplot(aes(x=wt, y=hp, color=mpg)) +
  geom_smooth(method="lm", se=FALSE,
              color=get_hurwitz_colors("light_gray")) +
  geom_point() +
  scale_color_hurwitz(discrete=FALSE, palette="classic") +
  labs(x="Weight (1000 lb)",
       y="Power (hp)",
       color="Fuel Efficiency (mpg)",
       title="Scatterplot of mtcars")
```
### The Goal

Let's assume that the goal of this is to create a model that predicts `mpg` based
on `weight` and `power`.


## Linear Regressions

### Multiple Linear Regressions {.tabset}

Keeping the goal in mind, one approach would be to create a simple linear regression.

$y = mx + b$

Where:

* $y$: `mpg`
* $m$: slope
* $b$: intercept
* $x$: independent variable (`weight` or `power`)

```{r lm, echo=FALSE}
lm_model <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```


Since we have 2 predictors (`weight` and `power`), then we could make 2 separate
linear regressions, and choose whichever one has a better fit to be our predictor.

```{r multiple_regressions, echo=FALSE}
wt_model <- lm_model %>% 
  fit(mpg ~ wt, data=train)

hp_model <- lm_model %>% 
  fit(mpg ~ hp, data=train)
```

#### Weight

```{r wt_model, echo=FALSE}
wt_rs <- round(summary(wt_model$fit)$r.squared, digits = 3)

train %>% ggplot(aes(x=wt, y=mpg)) +
  geom_smooth(method="lm", se=FALSE,
              color=get_hurwitz_colors("light_gray")) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(title="MPG vs Weight",
       subtitle=str_glue("r squared = {wt_rs}"),
       x="Weight (1000 lb)",
       y="Fuel Efficiency (mpg)")
```
#### Power

```{r hp_model, echo=FALSE}
hp_rs <- round(summary(hp_model$fit)$r.squared, digits = 3)

train %>% ggplot(aes(x=hp, y=mpg)) +
  geom_smooth(method="lm", se=FALSE,
              color=get_hurwitz_colors("light_gray")) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(title="MPG vs Power",
       subtitle=str_glue("r squared = {hp_rs}"),
       x="Power (hp)",
       y="Fuel Efficiency (mpg)")
```
### {-}

Between these two regressions (click the tabs), weight looks like it may be a better predictor than power based on $r^2$.

Wouldn't it be nice to have a model that creates a regression on both at the same time?

---

### Multilinear Regression

Multilinear regression (MLR) does exactly this. The formula in our case would be:
$y=m_1x_1 + m_2x_2 + b$

Where:

* $y$: `mpg`
* $x_1$: `weight`
* $m_1$: coefficient for `weight`
* $x_2$: `power`
* $m_2$: coefficient for `power`
* $b$: intercept

After performing this fit, here is the resulting MLR. 

```{r regression, echo=FALSE}
lm_fit <- lm_model %>% 
  fit(mpg ~ ., data=train)

m1 <- round(summary(lm_fit$fit)$coefficients["wt",1], 2)
m2 <- round(summary(lm_fit$fit)$coefficients["hp",1], 2)
b <- round(summary(lm_fit$fit)$coefficients["(Intercept)",1], 2)
rsq <- round(summary(lm_fit$fit)$r.squared, digits = 3)

print(str_glue("y = {m1}*weight + {m2}*hp + {b}"))
print(str_glue("r squared = {rsq}"))
```

We can see that this is a better model than either of the individual linear regressions.

But imagine if you have a very large number of predictors. There would have to be a term for each one. This is one of many reasons why we may want to perform dimension reduction. This will reduce the number of predictors so that the regression model is simplified.

## Projections

Although our dataset already has very few predictors (2), it can be used to illustrate the logic of dimension reduction very well.

So, lets go about reducing the number of predictors from 2 to 1.

---

### Correlated Predictors

Let's go back to a plot of our 2 predictors.

```{r corr_plot, echo=FALSE}
train %>% ggplot(aes(x=wt, y=hp)) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="Weight (1000 lb)",
       y="Power (hp)",
       title="Correlated Predictors")
```
It is quite apparent that these two predictors are correlated. As weight increases, power generally increases too.

Another way to put it is that if we know the weight, we know *something* about the power, and vice-versa. Of course the relationship isn't perfect -- we cannot predict power with perfect accuracy based on weight.

---

### Least Squares Regression

One way that we can see the correlation is with least-squares regression. This is what we generally mean when we say "linear regression", and is the method that was used in the previous section on MLR.

Least-squares regression finds the line that minimizes the magnitude of the errors. But is it important to note that it is defining errors based only on the $y$ values, i.e. the difference between the actual $y$ values and the predicted $y$ values. These errors can be visualized as vertical lines.

```{r pred_reg, echo=FALSE}
dr_model <- lm_model %>% 
  fit(hp ~ wt, data=train)

dr_int <- summary(dr_model$fit)$coefficients["(Intercept)",1]
dr_slope <- summary(dr_model$fit)$coefficients["wt",1]
```

```{r vert_bars, echo=FALSE}
train %>% ggplot(aes(x=wt, y=hp))+
  geom_segment(aes(xend=wt, yend=hp-dr_model$fit$residuals),
               colour=get_hurwitz_colors("viking")) +
  geom_abline(slope=dr_slope, intercept = dr_int,
              color=get_hurwitz_colors("light_gray"),
              size=1) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="Weight (1000 lb)",
       y="Power (hp)",
       title="Least-Squares Regression")
```


While the fact that least-squares regression is based on "vertical" errors may seem subtle, it can have certain implications, as we will see later.

---

### Regression Line as a New Axis?

Getting back to the objective: reducing our 2 predictors to 1.

This may not be an immediately obvious choice, but what if we converted our 2 predictors into 1 by taking the regression line to be a new axis.

Each point could then be represented by a single number: how far along the regression line that point lies.

To see how far along the regression line a point lies, we can draw a line segment that is perpendicular to the regression line, connecting the point to the regression line.

Before we try this with our dataset, let me illustrate the idea with some made up data.

```{r perps, echo=FALSE, fig.height=5, fig.width=5}
df_fake <- tibble(
  x = c(0.23, 0.71, 1.02, 1.28, 1.63, 1.87, 1.98,
        2.34, 2.8, 3.2, 3.56, 3.9, 4.5, 4.65),
  y = c(0.62, 0.89, 0.99, 1.95, 1.54, 2.07, 2.12,
        2.95, 2.2, 3.8, 3.6, 3.43, 4.1, 4.94)
)

fake_model <- lm_model %>% 
  fit(y ~ x, data=df_fake)

fake_int <- summary(fake_model$fit)$coefficients["(Intercept)",1]
fake_slope <- summary(fake_model$fit)$coefficients["x",1]

segments <- perp_segment_coord(df_fake$x, df_fake$y, fake_int, fake_slope) %>% as.data.frame()

df_fake %>% ggplot(aes(x=x, y=y)) +
  geom_segment(data=segments, 
               aes(x=x0, y=y0, xend=x1, yend=y1), 
               colour=get_hurwitz_colors("viking")) +
  geom_abline(slope=fake_slope, intercept = fake_int,
              color=get_hurwitz_colors("light_gray"),
              size=1) +
  geom_point()+
  xlim(c(0,5)) + ylim(c(0,5)) +
  labs(title="Fake Data with Regression Line",
       subtitle="Perpendicular lines show projection to new axis")
```
In this case, we could ignore our original 2 variables and instead use a simplified variable: the distance along the regression line where each point projects, as indicated by the blue lines. Since the original variables ($x$ and $y$) were correlated, this new variable still captures the information pretty well.

One thing to note though is that not all information is preserved. For instance the points at $(1.28, 1.95)$ and $(1.63, 1.54)$ project to nearly the same point on the regression line. So with just this 1 dimension, the difference between these points is greatly reduced.

With this ideal example in mind, let's do the same to our car data. Let's add perpendicular lines connecting the regression line and the points to illustrate the new axis.

```{r reg_axis, echo=FALSE}
reg_axis <- train %>% plot_perp() +
  labs(x="Weight (1000 lb)",
       y="Power (hp)",
       title="Regression Line as New Axis")

reg_axis
```
You may think, "Those are not perpendicular to the regression line!". They actually are, but because the scale on the 2 axes are so different, they don't look like it.

We can visualize the same plot with both axes to the same scale, but then we can't even see the lines. The magnitude of our vertical axis is just too much greater than our horizontal axis.

```{r rescaled_reg_axis, echo=FALSE, fig.height=5, fig.width=5}
max_scale <- 1.05*(max(train$hp, train$wt))

reg_axis +
  xlim(c(0,max_scale)) +
  ylim(c(0,max_scale)) +
  labs(title="Equally Scaled Regression")
```


### The Problem of Scale

Looking at the last plot, we can see that if we use this technique, then when we project the points to the regression line, the variable with smaller magnitude doesn't really seem to matter. Effectively, our new axis based on the regression line is pretty much the same as just using the `power` or $y$ axis.

The reality of this can be seen in this way. The regression line and vertical axis are pretty much parallel. So, when projecting to the regression line, it is about the same as projecting just to the vertical axis, and ignoring the other variable (`weight`).

So if we were to use this regression line as our new axis for dimension reduction, we would pretty much just be taking our new variable to be the `power` and ignore the `weight` variable.

---

### Rescaling {.tabset}

While the above commentary may make it seem like this approach is doomed, it actually shows something very different. Yes, it is true that this method is sensitive to the scale of the predictors, but the scale is actually quite arbitrary.

Let's look at the `weight` axis in the above plots. As noted, these are in units of 1000 lb. So a weight of 5 is actually 5000 lb, for example.

I can easily adjust my data so that the $x$ axis is in units of lb, which is completely valid. But this certainly affects the scale of the axis, and also our perception of the projection to the regression line.

#### Normal axis scaling

```{r, echo=FALSE}
rescaled <- train

rescaled$wt <- rescaled$wt * 1000

resc_plot <- rescaled %>% plot_perp() +
  labs(x="Weight (lb)",
       y="Power (hp)",
       title="Regression Line as New Axis, Rescaled Weight",
       subtitle="Lines are perpindicular to regression line")

resc_plot
```
#### Equal Axis Scales

```{r eq_axes, echo=FALSE, fig.height=5, fig.width=5}
max_scale_resc <- max(rescaled$wt, rescaled$hp)

resc_plot +
  xlim(c(0, max_scale_resc)) +
  ylim(c(0, max_scale_resc)) +
  labs(subtitle="",
       title="Rescaled Weight, Axes Scaled Equally")
```
### {-}

Now, instead of the lines that are perpendicular to the regression line looking horizontal, they look vertical. This is because the $x$ axis now has greater magnitude.

Additionally, the regression line is more parallel to the $x$ axis. So if we were to project onto this regression line, our new variable would mostly reflect the `weight` as a predictor.

*Takeaways:*

* The scaling of the axes is arbitrary
* Axis scales strongly affect how our predictor variables would be represented in a new variable based on projection to a regression line.

## Rescaling

The reason that when the fake data were projected to the new axis (regression line) worked so well is that I intentionally gave each variable a similar scale (0 to 5).

So, in order to make this proposed method of dimension reduction work, we need to ensure that each variable is equally scaled.We can rest assured that this is okay, since we have already demonstrated that scale is arbitrary ($1000 lb$ vs $lb$).

We could, for instance, pick a value for each variable by which we would divide each value. By choosing the right number for each, they could have the same scale. But how would you choose this value? Trial and error? Spray and pray?

There is actually a more robust method of rescaling. We can divide the values of each variable by that variable's standard deviation. If the values of each variable are least kind of normally distributed, they will have similar scale. Let's take a look at the results of this.

```{r norm, echo=FALSE, fig.height=5, fig.width=5}
norm <- train

norm$hp <- norm$hp / sd(norm$hp)
norm$wt <- norm$wt / sd(norm$wt)

norm_max <- norm %>% select(hp, wt) %>% max()

norm_plot <- norm %>% plot_perp() + 
  labs(x="Rescaled Weight",
       y="Rescaled Power",
       title="Rescaled Variables",
       subtitle="Variables divided by its standard deviation") +
  lims(x=c(0, norm_max), y=c(0, norm_max))

norm_plot
```


---

### Replotting

Now that the data are rescaled, the projections to the regression line are looking good!

Going back to the overall goal: we are trying to create a regression model to predict `mpg` based on a single variable that captures the information of the 2 original variables `weight` and `power`.

To do this, I will create a new variable which I will call `reg`, which is the distance along the regression line where each point projects. This can be seen as the where the blue lines touch the regression line in the plot above.

For now, I am doing this based on geometry (read pythagorean theorem).

So now, each point is represented by the single variable `reg`. We can plot this variable against `mpg` since this is the variable we want to predict based on our new single predictor variable, `reg`.

```{r rot, echo=FALSE}
df_rot <- train %>% plot_reg_coord()

rot_plot <- df_rot %>% ggplot(aes(x=x, y=mpg)) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="New variable, reg",
       y="Fuel Efficiency (mpg)",
       title="Fuel Efficiency vs New Variable")

rot_plot
```

We can now perform a linear regression on the above plot to see how well our new variable predicts `fuel efficiency`.

It is actually impressive! We have reduced the number of predictor variables, yet the correlation is the same as or stronger than:

* Linear regression based on `power` ($r^2=0.625$)
* Linear regression based on `weight` ($r^2=0.703$)
* Multilinear regression based on both `weight` and `hp` ($r^2=0.8$)

Of course, this is only a single observation, so we can't be sure if the difference is significant. Still, it is interesting.

```{r, echo=FALSE}
rot_fit <- lm_model %>% 
  fit(mpg ~ x, data=df_rot)

intercept <- summary(rot_fit$fit)$coefficients["(Intercept)",1]
slope <- summary(rot_fit$fit)$coefficients["x",1]

reg_rs <- round(summary(rot_fit$fit)$r.squared, digits = 3)
  
rot_plot1 <- df_rot %>% ggplot(aes(x=x, y=mpg)) +
  geom_abline(slope=slope, intercept=intercept,
                color=get_hurwitz_colors("light_gray"),
                size=1) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="New variable, reg",
       y="Fuel Efficiency (mpg)",
       title="Fuel Efficiency vs New Variable",
       subtitle=str_glue("r squared = {reg_rs}"))

rot_plot1
```
## Orthogonal Regression

### New Coordinate System

To review: our new variable `reg` is the projection of our data onto a linear
regression line of the re-scaled `weight` and `hp` variables. This regression
was performed using ordinary least-squares, which minimizes the errors in the
$y$ direction.

So far, we have only looked at the single new variable, represented by the
projection. But even in this new coordinate system, we can look at it as
having 2 dimensions:

* $x$: regression line axis
* $y$: axis perpendicular to regression line.

The new $y$ axis coordinate can be seen as the distance a point is from the new
$x$ axis, the regression line. These distances are shown in blue here.

```{r, echo=FALSE, fig.width=5, fig.height=5}
norm_plot
```

We can plot the same points in this new *rotated* coordinate system.

I will now refer to the regression axis as `reg1` and the perpendicular axis
as `reg2`.

```{r, echo=FALSE}
rot_plot <- df_rot %>% ggplot(aes(x=x, y=y)) +
  geom_hline(yintercept=0, color="lightgray", size=1) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="reg1",
       y="reg2",
       title="Rotated Coordinate System Plot")

rot_plot
```

---

### Persistent Correlation

It may (or may not) seem odd that these new variables are still correlated!

The strangeness may be more apparent when you think about why we took this
approach in the first place. We realized our variables were correlated,
so a projection to a regression line summarized both variables pretty well.
Since they're still correlated, we could do this same process again. And again..

Clearly, these variables are less correlated though
than the original variables (`weight` and `hp`). We would expect this to be the
case because *the regression on the original variables should have had points
distributed evenly above and below the regression line.* After all, isn't that
the whole idea of regression?

So given that last fact, why are they still correlated?

It comes down to how we performed the regression. As mentioned, we used
ordinary least-squares regression, which minimizes the vertical errors. By 
contrast, it did not minimize the real (perpendicular) distances from the
regression line.

This is problematic in a few ways: 

* Firstly, it means that this method is
dependent on which variable you put on which axis when performing regression.
This is undesirable, because you have to be cautious and choosy to get the "best"
results. Imagine brute-force comparing thousands of variable order combinations!
* Second, it means that our new variables are still (potentially)
correlated. This seems harmless, but it opens the idea of, why not do the method
again. When do you decide to stop?

---

### Orthogonal Regression {.tabset}

One way around this is to perform a different kind of linear regression.
Orthogonal (*a.k.a* total least-square or Deming) regression minimizes the
errors orthogonal (perpendicular) to the regression line.

Below, I have shown both types of regression. The errors which are minimized for
each method are shown in blue. I have also performed the projection to the
different regression lines, and plotted the data in the new coordinate system.

#### Regressions

It is apparent that the regression lines are very different in each case.

```{r, echo=FALSE, fig.width=10, fig.height=4}
dem_fit <- Deming(norm$wt, norm$hp)

norm_fit <- lm_model %>% 
  fit(hp ~ wt, data=norm)

demi_intercept <- dem_fit["Intercept"]
demi_slope <- dem_fit["Slope"]

intercept <- summary(norm_fit$fit)$coefficients["(Intercept)",1]
slope <- summary(norm_fit$fit)$coefficients["wt",1]

demi_segments <- perp_segment_coord(norm$wt, norm$hp, demi_intercept, demi_slope) %>% 
  as.data.frame()

plot <- norm %>% ggplot(aes(x=wt, y=hp)) +
  labs(x="Rescaled Weight",
       y="Rescaled Power")

ord_plot <- plot +
  geom_abline(slope=slope, intercept=intercept,
              color=get_hurwitz_colors("light_gray"),
              size=1) +
  geom_segment(aes(xend=wt, yend=hp-norm_fit$fit$residuals),
               colour=get_hurwitz_colors("viking")) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(title="Ordinary Regression")

deming_plot <- plot +
  geom_abline(slope=demi_slope, intercept=demi_intercept,
              color=get_hurwitz_colors("light_gray"),
              size=1) +
  geom_segment(data=demi_segments, 
               aes(x=x0, y=y0, xend=x1, yend=y1), 
               colour=get_hurwitz_colors("viking")) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(title="Orthogonal Regression",
       y="")

grid.arrange(ord_plot, deming_plot, ncol=2)

```

#### New Coordinates

We can see that the new coordinates based on orthogonal regression are 
uncorrelated as indicated by the horizontal regression line
and $r^2=0$. The variables based on ordinary regression are only weakly correlated, but the regression line is obviously not horizontal.

```{r, echo=FALSE, fig.width=10, fig.height=4}
rot_fit2 <- lm_model %>% 
  fit(y ~ x, data=df_rot)

intercept <- summary(rot_fit2$fit)$coefficients["(Intercept)",1]
slope <- summary(rot_fit2$fit)$coefficients["x",1]

rot_rs <- round(summary(rot_fit2$fit)$r.squared, digits = 3)

rot_plot <- df_rot %>% ggplot(aes(x=x, y=y)) +
  geom_abline(slope=slope, intercept=intercept,
              color=get_hurwitz_colors("light_gray"),
              size=1, alpha=0.5) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="reg1",
       y="reg2",
       title="Rotated Ordinary Regression",
       subtitle=str_glue("r square = {rot_rs}"))

demi_rot_df <- norm

demi_rot_df <- demi_rot_df %>% add_column(
  x = sqrt((demi_segments$x1)^2 + (demi_segments$y1 - demi_intercept)^2),
  y = sqrt(
    (demi_segments$x1 - demi_segments$x0)^2 +
      (demi_segments$y1 - demi_segments$y0)^2))

demi_rot_df <- demi_rot_df %>% mutate(x = case_when(
  demi_segments$x1 < 0 ~ -x,
  demi_segments$x1 >= 0 ~ x
)) %>% 
  mutate(y = case_when(
    demi_segments$y0 < demi_segments$y1 ~ -y,
    demi_segments$y0 >= demi_segments$y1 ~ y
  ))

demi_rot_fit <- lm_model %>% 
  fit(y ~ x, data=demi_rot_df)

demi_rot_int <- summary(demi_rot_fit$fit)$coefficients["(Intercept)",1]
demi_rot_slope <- summary(demi_rot_fit$fit)$coefficients["x",1]

demi_rot_rs <- round(summary(demi_rot_fit$fit)$r.squared, digits = 3)

demi_rot_plot <- demi_rot_df %>% ggplot(aes(x=x, y=y)) +
  geom_abline(slope=demi_rot_slope, intercept=demi_rot_int,
            color=get_hurwitz_colors("light_gray"),
            size=1, alpha=0.5) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="ortho1",
       y="ortho2",
       title="Rotated Orthogonal Regression",
       subtitle=str_glue("r square = {demi_rot_rs}"))

grid.arrange(rot_plot, demi_rot_plot, ncol=2)
```

### {-}

---

### Checking Predictor

Now that we have been able to find a better regression model to project to, let's
see how this new variable predicts fuel efficiency. For the regressions shown below ordinary least squares regression is used at this point, since we are trying to predict fuel efficiency based on our new variable. *ortho* and *ordinary* in the titles indicate the regression method used for generating the independent variable.

The correlation seems to be about the same as the variable made from ordinary regression projection. While there is no obvious improvement, all the previous commentary on the benefit of orthogonal regression still holds.

```{r, echo=FALSE, fig.width=10, fig.height=4}
demi_reg <- lm_model %>% 
  fit(mpg ~ x, data=demi_rot_df)

demi_reg_int <- summary(demi_reg$fit)$coefficients["(Intercept)",1]
demi_reg_slope <- summary(demi_reg$fit)$coefficients["x",1]

demi_reg_rs <- round(summary(demi_reg$fit)$r.squared, digits = 3)

demi_reg_plot <- demi_rot_df %>% ggplot(aes(x=x, y=mpg)) +
  geom_abline(slope=demi_reg_slope, intercept=demi_reg_int,
              color=get_hurwitz_colors("light_gray"),
              size=1) +
  geom_point(color=get_hurwitz_colors("red")) +
  labs(x="ortho1",
       y="",
       title="Ortho Projected Regression",
       subtitle=str_glue("r square = {demi_reg_rs}"))

rot_plot1 <- rot_plot1 +
  labs(title="Ordinary Projected Regression",
       x="reg1")

grid.arrange(rot_plot1, demi_reg_plot, ncol=2)
```

## Eigenvectors

So far we have explored 2 methods of finding a new axis to which we can project:

* Ordinary least-squares regression
* Orthogonal leas-squares regression

We began with regressions since these intuitively follow directions of variance. However, a more efficient method is to find the eigenvectors of the covariance matrix between the variables. These eigenvectors will serve as the new axes for data projection, similar to how the regression lines were used.

---

Covariance Matrix

```{r}
cov_df <- cor(norm$wt, norm$hp)

round(cov_df, 2)
```
```{r}
eig_df <- eigen(cov_df)

eig_df
```
```{r}
eig_t <- t(eig_df$vectors)

norm_t <- norm %>% select(wt, hp) %>% t()

eig_proj <- eig_t %*% norm_t

norm_eig <- t(norm_t)
```
```{r}
pca_fit <- norm %>% 
  select(-mpg) %>% 
  prcomp(scale=FALSE, center=FALSE)

pca_reg_plot <- demi_rot_df %>% ggplot(aes(y=x)) +
  # geom_abline(slope=demi_reg_slope, intercept=demi_reg_int,
  #             color=get_hurwitz_colors("light_gray"),
  #             size=1) +
  geom_point(aes(x=pca_fit$x[,1]),
    color=get_hurwitz_colors("red")) +
  labs(x="PC1",
       y="mpg",
       title="PCA Projected Regression",
       subtitle=str_glue("r square = {}"))

pca_reg_plot
```

